### 2014-05-02 stat246/hw4.tex
s4

Now, differentiating wrt $\sigma'$ and setting to $0$ gives
\begin{align*}
    0&=\sum_{i=1}^N\int_\rn q_i(y)\left(\frac{(y-\mu')^2}{\sigma'^3}-\frac{1}{\sigma'}\right)\,dy\\
    N&=\frac{1}{\sigma'^2}\sum_{i=1}^N\int_\rn q_i(y)(y-\mu')^2\,dy\\
    \sigma'^2&=\frac{1}{N}\sum_{i=1}^NE_{Y|X=x_i}(Y-\mu')^2\\
\end{align*}

The expectation inside the sum expands to $\var(Y)+E(Y)^2-2\mu' E(Y)+\mu'^2$, so the final expression for $\sigma'^2$ is
\[\mu'^2+\frac{1}{N}\sum_{i=1}^N\var(Y|x_i)+E(Y|x_i)^2-2\mu'E(Y|x_i)\]
\ssn{e}
\end{document}

>>>>

Now, differentiating wrt $\sigma'$ and setting to $0$ gives
\begin{align*}
    0&=\sum_{i=1}^N\int_\rn q_i(y)\left(\frac{(y-\mu')^2}{\sigma'^3}-\frac{1}{\sigma'}\right)\,dy\\
     &=\sum_{i=1}^N\int_\rn q_i(y)\left(\frac{(y-\mu')^2}{\sigma'^2}-1\right)\,dy\\
    N&=\frac{1}{\sigma'^2}\sum_{i=1}^N\int_\rn q_i(y)(y-\mu')^2\,dy\\
    \sigma'^2&=\frac{1}{N}\sum_{i=1}^NE_{Y|X=x_i}(Y-\mu')^2\\
\end{align*}

The expectation inside the sum expands to $\var(Y)+E(Y)^2-2\mu' E(Y)+\mu'^2$, so the final expression for $\sigma'^2$ is
\[\mu'^2+\frac{1}{N}\sum_{i=1}^N\var(Y|x_i)+E(Y|x_i)^2-2\mu'E(Y|x_i)\]
where the dependence on the previous $\sigma$ comes in from the conditional expectations and variance
\ssn{e}
If we let $\mu$ be the sample mean of the $X_i$ in the expression for $\mu'$ above, then the sum cancels out to zero and we are left with $\mu'=\mu$, which means that it's a fixed point. This is the only fixed point, since the equation obtained by setting $\mu'=\mu$ is linear, which means that there's only one solution.
\ssn{f}
If we set $\mu=\mu'=\h{\mu}_{ML}$, then our expression for $\sigma'^2$ becomes
\[\h{\mu}^2+\left(1-\frac{\sigma^2}{\sigma^2+\sigma_1^2}\right)\sigma^2+\frac{1}{N}\sum_{i=1}^N\left(\h{\mu}+\frac{\sigma^2}{\sigma^2+\sigma_1^2}(x_i-\h{\mu})\right)^2-\frac{2\h{\mu}}{N}\sum_{i=1}^N\h{\mu}+\frac{\sigma^2}{\sigma^2+\sigma_1^2}(x_i-\h{\mu})\]
The term with the unadorned sum of the $x_i-\h{\mu}$ sums out to zero, so canceling that and expanding the square leaves us with
\[\left(1-\frac{\sigma^2}{\sigma^2+\sigma_1^2}\right)\sigma^2+\frac{1}{N}\sum_{i=1}^N2\h{\mu}\frac{\sigma^2}{\sigma^2+\sigma_1^2}(x_i-\h{\mu})+\frac{1}{N}\sum_{i=1}^N\left(\frac{\sigma^2}{\sigma^2+\sigma_1^2}(x_i-\h{\mu})\right)^2\]
The middle term again disappears, and if we're assuming that $\sigma=\h{\sigma}_{ML}$, then we can rewrite the last term:
\[\sigma^2-\frac{\sigma^4}{\sigma^2+\sigma_1^2}+\frac{\sigma^4(\sigma^2+\sigma_1^2)}{(\sigma^2+\sigma_1^2)^2}=\sigma^2\]
so we have that the ML estimate is a fixed point.
\subsection*{3}
\ssn{a}
Let $Y_1,\ldots, Y_N$ be the hidden class variables. The complete-data log-likelihood is then
\[\sum_{i=1}^N\left(\log \pi_{y_i}+\sum_{j=1}^dx_{i,j}\log p_{j,y_i}+(1-x_{i,j})\log(1-p_{j, y_i})\right)\]
Taking the expectation wrt $Y$ given previous parameters and the data yields
\[\sum_{i=1}^N\sum_{m=1}^Mw_{i,m}\left(\log\pi_m+\sum_{j=1}^dx_{i,j}\log p_{j,m}+(1-x_{i,j})\log(1-p_{j,m})\right)\]
The ML estimate of the $\pi_m$ is, as usual, $\frac{1}{N}\sum_{i=1}^Nw_{i,m}$. Differentiating wrt each $p_{j,m}$, we have
\[\sum_{i=1}^Nw_{i,m}\left(\frac{x_{i,j}}{p_{j,m}}-\frac{1-x_{i,j}}{1-p_{j,m}}\right)\]
Setting this equal to zero produces 
\[p_{j,m}=\frac{\sum_{i=1}^Nw_{i,m}x_{i, j}}{\sum_{i=1}^N w_{i,m}}\]
\ssn{b}
We want conjugate priors for $\pi$ and $p$. Since the conjugate priors for the Bernoulli and the multinomial are the beta and the Dirichlet, respectively, we use $Md$ independent beta densities for the $p_{j,m}$ and a Dirichlet for all the $\pi$s. For simplicity's sake, initialize all the beta distributions with the same hyperparameter $\alpha$. Let the initial hyperparameters for the Dirichlet be $\zeta_m$.

The quantity to maximize is now $l(\theta|X,Y)+\log p(\theta)$, which in this case is
\begin{gather*}
    \sum_{i=1}^N\sum_{m=1}^Mw_{i,m}\left(\log\pi_m+\sum_{j=1}^dx_{i,j}\log p_{j,m}+(1-x_{i,j})\log(1-p_{j,m})\right)\\
    +\sum_{i=1}^N\sum_{m=1}^M\sum_{j=1}^d\frac{1}{N}\left((\alpha-1)\log p_{j,m}+(\alpha-1)\log(1-p_{j,m})\right)\\
    +\frac{1}{N}\sum_{i=1}^N\sum_{m=1}^M(\zeta_m-1)\log\pi_m
\end{gather*}
after removing constants irrelevant to maximization. Folding the sums in gives
\[\sum_{i=1}^N\sum_{m=1}^M\left(w_{i,m}+\frac{1}{N}(\zeta_m-1)\right)\log\pi_m+\sum_{i=1}^N\sum_{m=1}^Mw_{i,m}\sum_{j=1}^d\left(x_{i,j}+\frac{1}{N}(\alpha-1)\right)\log p_{j,m}+\left(1-x_{i,j}+\frac{1}{N}(\alpha-1)\right)\log(1-p_{j,m})\]
Having the same form as the thing we maximized already, we find that the new equations for $\pi_m$ and $p_{j,m}$ are
\begin{align*}
    \pi_m&=\frac{1}{N}\sum_{i=1}^N\left(w_{i,m}+\frac{1}{N}(\zeta_m-1)\right)\\
    p_{j,m}&=\frac{\sum_{i=1}^Nw_{i,m}\left(x_{i,j}+\frac{1}{N}(\alpha-1)\right)}{\left(\frac{2}{N}(\alpha-1)+1\right)\sum_{i=1}^Nw_{i,m}}
\end{align*}
\end{document}
